---
title: "Big Data Republic"
subtitle: "Tanzanian Ministry Water pumps"
author: "Angeliki Kalamara, PhD"
#institute: ""
date: "08/10/2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{css echo=FALSE}
.darkblue {
  color: #1A5B8B;
}

.remark-slide-content > h1 {   font-size: 33px;   margin-top: -80px; }
```

```{r include=FALSE, message=FALSE}
library(dplyr)
library(highcharter)

df_wp_set=read.csv('water_pump_set.csv',header = T)
df_wp_labels=read.csv('water_pump_labels.csv',header=T)

df_merged = merge(df_wp_set,df_wp_labels,by='id')
```


# Problem statement

Tanzanian Ministry of Water is interested in improving the maintenance of the Tanzanian water pumps. 

**The objective is to develop a reproducible model that can predict the failure of the pumps in the future.**

--

*Suggestion for deployment of the model acts as bonus!*

---
# Dataset

.pull-left[
- 59,400 rows (unique pumps)
- 41 column (features)
- **'status_group'** target variable
    - 'functional'
    - 'functional, needs repair'
    - 'non-functional'
- 9 numerical features (id dropped) and 31 categorical features (target variable included).
]

.pull-right[
```{r echo=FALSE, message=FALSE, out.width = '2000px', out.height = '400px'}
df_status_pump = df_merged %>% group_by(status_group) %>% summarise(n=n()) %>% as.data.frame()
df_status_pump = df_status_pump %>% mutate(perc = paste0(round(n/nrow(df_merged) * 100,2),'%'))

library(ggplot2)
ggplot(df_status_pump, aes(x=status_group, y=n, fill = status_group)) +
    geom_bar(stat="identity", width=0.6)+
    geom_text(aes(label=paste0(n,' (',perc,')')), 
              position=position_dodge(width=0.2), hjust=0.5, vjust=-0.5, size=5) + 
    scale_fill_manual(values=c('#42A0E8','#F2A808', 'gray')) + ggtitle('Pump status')+
    labs(fill='Pump Status') +
    theme_minimal() + 
    expand_limits(y=c(0, 33000)) +
    theme(
        title=element_text(face='bold'),
        legend.position = 'top',
        legend.text = element_text(size=13),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.text.x=element_text(size=15,angle=45,hjust=0.9,vjust=0.9,face='bold'),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```
]

---
# Data Preprocessing (1): Drop features

- Transform 'date_recorded' to mumeric (months passed untill 01/-1-2021) and drop it
- Drop 'recorded_by' (only 'GeoData Consultants Ltd') hence not very informative for the model
- Drop 'wpt_name' (waterpoint name) and 'num_private' as for the latter feature no explanation is provided in the data set legend and contains only numbers

---

# Data Preprocessing (2): Group features

Group features by dropping the ones with similar values. Many categorical feature are very similar to each other; in this case we keep the most informative afer visual inspection.

A. Extraction type ('extraction_type', 'extraction_type_group', **'extraction_type_class'**): latter is more coincise

B. Quantity (**'quantity'**, 'quantity_group'): exactly the same

C. Payment ('payment', **'payment_type'**): exactly the same

D. Waterpoint type ('waterpoint_type', **'waterpoint_type_group'**): better/less redundant categorization (communal standpipe + communal standpipe multiple together in one category)

E. Management (**'management'**, 'management_group'): latter appears very generic; first provides a little more info; management of the waterpoint is critical as it can directly contribute to failure.

F. Source (**'source'**, 'source_type', 'source_class'): could be important feature for predicting pump failure; 'source' was kept as sufficiently descriptive.


<i><span class="darkblue">The features in <b>bold</b> are the ones kept to be used to build the predictive model.</span></i>

---

#  Data Preprocessing (3): Handle 0's

Handle O's in the dataset:

- Construction year; replace 0 values with median construction year
- Population - has lots of zeros (aka NAs); replace with median polulation

---

# Data Preprocessing (4): Handle NULL/NA values

```{r echo=FALSE}
xx = sapply(colnames(df_merged), function(x) {
    which(df_merged[,x] == '') %>% length
})

xx = xx[xx!=0] %>% sort
df_null = data.frame(feature = names(xx), num_null = as.numeric(xx))

knitr::kable(df_null, format = 'html')
```

Drop these features
---

# Spread of water pumps in Tanzania

```{r echo=FALSE, message=FALSE}
library(viridisLite)
library(highcharter)
library(dplyr)
df_pump_per_region = df_merged %>% group_by(region) %>% summarise(total=n()) %>% as.data.frame()

mapdata <- download_map_data("https://code.highcharts.com/mapdata/countries/tz/tz-all.js")

highchart() %>%
  hc_add_series_map(mapdata, df_pump_per_region, name = "Water Pumps",
                    value = "total", joinBy = c("woe-name", "region"),
                    dataLabels = list(enabled = TRUE, format = '{point.properties.name}'))
```


